<!-- R Commander Markdown Template -->

Title: Paper-Airplane-Flight-Analysis
=======================

### author: “Siamak Goudarzi”

### date: “2026-01-09”

### output: pdf_document

### `r as.character(Sys.Date())`

```{r echo=FALSE}
# include this code chunk as-is to set options
knitr::opts_chunk$set(comment=NA, prompt=TRUE, out.width=750, fig.height=8, fig.width=8)
library(Rcmdr)
library(car)
library(RcmdrMisc)
```

```{r}
DS_Aufgabe5 <- 
  readXL("5_Technischer Vergleich mit Konkurrenz.xlsx",
   rownames=FALSE, header=TRUE, na="", sheet="Daten", stringsAsFactors=TRUE)
```


```{r}
library(abind, pos=17)
```


```{r}
library(e1071, pos=18)
```



### Zusammenfassungen für numerische Variablen: DS_Aufgabe5
```{r}
numSummary(DS_Aufgabe5[,c("Konkurrent.A", "Konkurrent.B", "Konkurrent.C", 
  "Konkurrent.D", "Unser.Modell"), drop=FALSE], statistics=c("mean", "sd", 
  "IQR", "quantiles"), quantiles=c(0,.25,.5,.75,1))
```

```{r}
Aufgabe5_StackedData <- stack(DS_Aufgabe5[, c("Konkurrent.A","Konkurrent.B",
  "Konkurrent.C","Konkurrent.D","Unser.Modell")])
names(Aufgabe5_StackedData) <- c("variable", "factor")
```


### Levenes Test: variable ~ factor
```{r}
Tapply(variable ~ factor, var, na.action=na.omit, 
  data=Aufgabe5_StackedData) # variances by group
leveneTest(variable ~ factor, data=Aufgabe5_StackedData, center="median")
```


### Quantile-Comparison Plot: variable
```{r}
with(Aufgabe5_StackedData, qqPlot(variable, dist="norm", id=list(method="y",
   n=2, labels=rownames(Aufgabe5_StackedData)), groups=factor))
```


### Test auf Normalverteilung: variable ~ factor
```{r}
normalityTest(variable ~ factor, test="shapiro.test", 
  data=Aufgabe5_StackedData)
```

```{r}
library(mvtnorm, pos=19)
```


```{r}
library(survival, pos=19)
```


```{r}
library(MASS, pos=19)
```


```{r}
library(TH.data, pos=19)
```


```{r}
library(multcomp, pos=19)
```



### Ein-Weg-Varianzanalyse: variable ~ factor
```{r}
AnovaModel.1 <- aov(variable ~ factor, data=Aufgabe5_StackedData)
summary(AnovaModel.1)
with(Aufgabe5_StackedData, numSummary(variable, groups=factor, 
  statistics=c("mean", "sd")))
local({
  .Pairs <- glht(AnovaModel.1, linfct = mcp(factor = "Tukey"))
  print(summary(.Pairs)) # pairwise tests
  print(confint(.Pairs, level=0.95)) # confidence intervals
  print(cld(.Pairs, level=0.05)) # compact letter display
  old.oma <- par(oma=c(0, 5, 0, 0))
  plot(confint(.Pairs))
  par(old.oma)
})
```


### Plot of Means: variable by factor
```{r}
with(Aufgabe5_StackedData, plotMeans(variable, factor, error.bars="se", 
  connect=TRUE))
```


### Boxplot: variable ~ factor
```{r}
Boxplot(variable ~ factor, data=Aufgabe5_StackedData, id=list(method="y"))
```

